Less Funding, More Lecturers, and Fewer Professors: \\ \vspace{0.1cm} Stagnating State Funding for Higher Education and its Effect on Faculty

### 2024-09-27 text/sections/01-introduction.tex
This is all done.

I provide the first causal evidence on the impact of state funding on faculty composition and wages.
In a contemporaneous working paper, \cite{hinrichs2022state} use the same national-level variation in university funding to analyse effects on faculty and wages, coming to different conclusions.\footnote{
    An anonymous reviewer made me aware of a similar and contemporary working paper, \cite{hinrichs2022state}; this work analyses a shorter time-period than this paper, and only focuses on national variation.
    \autoref{sec:recession-analysis} replicates their key findings, and shows how their focusing on only the 2012--2019 years explains differences in conclusions.
}
The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,webber2017state}.
For faculty outcomes, descriptive work measures how US faculty salaries are correlated with other factors \citep{hilmer2020labor}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
Other papers measure university decision--making after endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
These papers measure changes in state funding and university revenues, but do not measure effects of state funding cuts on faculty composition or salaries, or other outcomes related to university instruction.
My results provide evidence that the long-term trends in higher education funding and faculty outcomes are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.

>>>>

I provide the first causal evidence on the impact of state funding on faculty composition and wages.
In a contemporaneous working paper, \cite{hinrichs2022state} use the same national-level variation in university funding to analyse effects on faculty and wages, coming to different conclusions for the national-level analysis.\footnote{
    An anonymous reviewer made me aware of a similar and contemporary working paper, \cite{hinrichs2022state}.
    My findings are consistent with theirs when restricting to years 2012--2019, but show a substitution towards lecturers when expanding to the entire time period 1990--2021.
    See \autoref{sec:recession-analysis} for this replication.
}
The closest related research on higher education funding examined how university spending affected graduation rates and levels of student debt \citep{NBERw23736,NBERw27885}, and university finances \citep{miller2022making,bound2019public,webber2017state}.
For faculty outcomes, descriptive work measures how US faculty salaries are correlated with other factors \citep{hilmer2020labor}, and applied theory work models university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
Other papers measure university decision--making after endowment shocks \citep{brown2014endowment}, and trends in the academic job market following the 2008 recession \citep{turner2014impact}.
These papers measure changes in state funding and university revenues, but do not measure effects of state funding cuts on faculty composition or salaries, or other outcomes related to university instruction.
My results provide evidence that the long-term trends in higher education funding and faculty outcomes are causally related, contributing to the literature on faculty outcomes \citep{ehrenberg2003studying} and trends in US higher education and funding \citep{hoxby2009changing,ehrenberg2012american}.
